---
title: 'Homework #9'
author: "Advay Vyas"
date: 4/21/25
output:
  pdf_document:
    toc: true
urlcolor: blue
linkcolor: red
---

```{r global_options, echo=FALSE}
knitr::opts_chunk$set(fig.height=4, fig.width=6, fig.align = "center", warning=FALSE, echo=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60))
```

------------------------------------------------------------------------

```{r, results='hide', warning=FALSE, message=FALSE}
# loading libraries
library(tidyverse)
library(ggplot2)
library(lubridate)
library(sas7bdat)
library(rvest)
library(stringr)
library(boot)
library(mosaic)
library(MatchIt)
library(moderndive)
```

# Introduction
I'm Advay Vyas, EID: av37899, and this is my submission for SDS 315 Statistical Thinking Homework #9. The GitHub repository for my code is at this [link](https://github.com/advayvyas/SDSHW9).

\newpage

# Problem 1
```{r}
solder = read.csv("solder.csv")
ggsolder = ggplot(solder)
```

## Part A
```{r, fig.height = 3}
ggsolder + geom_boxplot(aes(y=skips, fill=Opening)) + facet_wrap(~Opening) + 
  labs(title = "Distribution of Solder Skips by Solder Gun Opening Size", y = "Skips") +
   theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())
```
From this plot, we can easily see that **the size of the opening positively affects the amount of solder skips** by examining the faceted box plots' vertical placement.

```{r, fig.height = 3}
ggsolder + geom_boxplot(aes(y=skips, fill = Solder)) + facet_wrap(~Solder) + 
  labs(title = "Distribution of Solder Skips by Alloy Thickness", y = "Skips") +
   theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())
```
From this plot, we can see that **alloy thickness has an effect on the amount of solder skips**, with thinner alloys contributing to usually higher amounts of solder skips.

## Part B
```{r}
solder_model = lm(skips ~ Solder + Opening + Solder:Opening, data = solder)
```

```{r, cache=TRUE}
results_solder = do(10000) * {
  model = lm(skips ~ Solder + Opening + Solder:Opening, data = mosaic::resample(solder))
  y_int = as.numeric(coef(model)[1])
  solder_t = as.numeric(coef(model)[2])
  opening_m = as.numeric(coef(model)[3])
  opening_s = as.numeric(coef(model)[4])
  solder_thin_m = as.numeric(coef(model)[5])
  solder_thin_s = as.numeric(coef(model)[6])
  data.frame(y_intercept = y_int,
             solder_thin = solder_t,
             opening_med = opening_m,
             opening_small = opening_s,
             solder_thin_med = solder_thin_m,
             solder_thin_small = solder_thin_s)
}
```

```{r, message=FALSE}
# y-intercept
ci_y_intercept <- confint(results_solder$y_intercept, level = 0.95)
est_y_intercept <- mean(results_solder$y_intercept)

# solder_thin
ci_solder_thin <- confint(results_solder$solder_thin, level = 0.95)
est_solder_thin <- mean(results_solder$solder_thin)

# opening_med
ci_opening_med <- confint(results_solder$opening_med, level = 0.95)
est_opening_med <- mean(results_solder$opening_med)

# opening_small
ci_opening_small <- confint(results_solder$opening_small, level = 0.95)
est_opening_small <- mean(results_solder$opening_small)

# solder_thin_med
ci_solder_thin_med <- confint(results_solder$solder_thin_med, level = 0.95)
est_solder_thin_med <- mean(results_solder$solder_thin_med)

# solder_thin_small
ci_solder_thin_small <- confint(results_solder$solder_thin_small, level = 0.95)
est_solder_thin_small <- mean(results_solder$solder_thin_small)

library(tibble)

results_table <- tibble(
  Condition = c(
    "Baseline", 
    "SolderThin", "OpeningMid", "OpeningSmall", 
    "SolderThin:OpeningMid", "SolderThin:OpeningSmall"
  ),
  Estimate = round(c(est_y_intercept,
    est_solder_thin, est_opening_med, est_opening_small, 
    est_solder_thin_med, est_solder_thin_small
  ), 3),
  CI_Lower = round(c(ci_y_intercept[1,1],
    ci_solder_thin[1, 1], ci_opening_med[1, 1], ci_opening_small[1, 1], 
    ci_solder_thin_med[1, 1], ci_solder_thin_small[1, 1]
  ), 5),
  CI_Upper = round(c(ci_y_intercept[1,2],
    ci_solder_thin[1, 2], ci_opening_med[1, 2], ci_opening_small[1, 2], 
    ci_solder_thin_med[1, 2], ci_solder_thin_small[1, 2]
  ), 5)
)

knitr::kable(results_table, caption = "Coefficients of the fitted linear regression model")
```

## Part C
If we round every coefficient to 0, we now create the equation below representing the amount of skips as $\hat{y}$ and the corresponding effect variables and y-intercept.
\begin{align*}
\hat{y} ={}&\ 0 
+ 2 \cdot \text{SolderThin} 
+ 2 \cdot \text{OpeningMid} 
+ 5 \cdot \text{OpeningSmall} \\
& - 1 \cdot \text{SolderThin} \cdot \text{OpeningMid} 
+ 10 \cdot \text{SolderThin} \cdot \text{OpeningSmall}
\end{align*}

* The y-intercept indicates that the predicted skips for SolderThick & OpeningLarge is 0 (the baseline)
* The effect of the SolderThin variable is 2, so you add two skips when you have SolderThin
* Similarly, this means the effect of having the OpeningMid variable raises the prediction by 2
* Having OpeningSmall raises the predicted amount of skips by 5
* Having both SolderThin and OpeningMid together reduces the predicted amount of skips by 1
* Having both SolderThin and OpeningSmall greatly increases the predicted amount of skips by 10


## Part D

We first have to tabulate every possible combination. For SolderThin, we immediately have 2 for the expected amount of skips. Considering small, medium, and large openings, we have that OpeningMid adds 2 and OpeningSmall adds 5 so 7, 4, 2 respectively. Considering the interactions, SolderThin and OpeningMid reduces the predicted skips by 1 and SolderThin and OpeningSmall increases the predicted skips by 10 resulting in 17, 3, 2 for SolderThin and OpeningSmall, OpeningMed, OpeningLarge respectively. 

The calculations become much simpler for SolderThick due to no interaction variables. We simply add the effect variable to 0 since SolderThick has "no effect." With OpeningSmall (+5), OpeningMid (+2), and OpeningLarge (+0 b/c no effect variable and it is the baseline) and we get 5, 2, 0 respectively. The results are tabulated in the table below.

```{r}
combos = data.frame (
  Solder = c("Thin", "Thin", "Thin", "Thick", "Thick", "Thick"),
  Opening = c("S", "M", "L", "S", "M", "L"),
  Skips = c(17,3,2,5,2,0)
)
 
knitr::kable(combos, caption = "Combinations of Solder Thicknesses and Solder Gun Openings and Predicted Skips")
```

Since soldering skips are manufacturing defects, I'd like to recommend the combination with the least amount of defects which is **SolderThick and OpeningLarge**. This trend is also seen in the earlier plots, as a thick solder and a large solder gun opening each had the least average skips in their categories.

# Problem 2
```{r}
groceries = read.csv("groceries.csv")
```

## Part A
```{r, fig.width = 8, fig.height = 4}
avg_price_store = groceries %>% group_by(Store) %>% 
  summarize(avg_price = mean(Price, na.rm = TRUE))

ggplot(avg_price_store) + geom_col(aes(x = reorder(Store, avg_price), y = avg_price), col = "black", fill = "skyblue") + 
  coord_flip() + labs(title = "Average Price of Products by Store", y = "Price (in dollars)", x = "")
```
This graph above shows the average price of a arbitrary product by store.

## Part B
```{r, fig.height = 6, fig.width = 10}
product_by_store = groceries %>% group_by(Product) %>% summarize(freq = n())

ggplot(product_by_store) + geom_col(aes(x = reorder(Product, freq), y= freq), col = "black", fill = "skyblue") + 
  coord_flip() +
  labs(title = "Products by Number of Stores Selling", x = "Product", y = "Stores")
```
This graph above shows the amount of stores that carry/sell each of these products.

## Part C
```{r, cache = TRUE}
results_type = do(10000) * {
  type_model = lm(Price ~ Product + Type, data = mosaic::resample(groceries))
  grocery_i = as.numeric(coef(type_model)["TypeGrocery"])
  data.frame(grocery_increase = grocery_i)
}
```

```{r, message = FALSE}
ci_grocery_store = confint(results_type$grocery_increase, level = 0.95)
est_grocery_store = mean(results_type$grocery_increase)
```

Compared with ordinary grocery stores (like Albertsons, HEB, or Krogers), **convenience stores charge somewhere between $0.45 and $0.88 more for the same product**.

## Part D
```{r}
store_model = lm(Price ~ Product + Store, data = groceries)
stores = coef(store_model) %>% round(5)
```

The two stores that charge the lowest prices for the same product are **Walmart (-0.99) and Kroger Fresh Fare (-0.90)**.
The two stores that charge the highest prices for the same product are **Whole Foods (+0.36) and Wheatsville Food Co-Op (+0.29)**.

## Part E
The coefficient for H-E-B is -0.65 while the coefficient for Central Market is -0.57, giving a difference of about 8 cents. If we compare this to other stores like the "cheap" Walmart & Kroger, we get about 9 cents. On the other hand, if we compare to the "expensive" Whole Fools & Wheatsville, those two stores have a difference of 7 cents. Even in close categories or "sides", we see that the differences are around 8 cents; from stores like Kroger to Whole Foods, the difference would easily exceed a dollar. Therefore, we can likely say that **Central Market charges similar prices to H-E-B for the same product**.

## Part F
```{r}
groceries$Income10K = groceries$Income / 10000
income_model = lm(Price ~ Product + Income10K, data = groceries)
```
**Consumers in poorer ZIP codes usually pay slightly more for a product on average** because Income10K had a coefficient of `r round(coef(income_model)["Income10K"][[1]], 3)` which indicates that for every $10,000 increase in average income for the ZIP code, the average price of products falls by a cent and a half.

The effect of this variable isn't terribly important as shown in the following sentence. A one-standard deviation increase in the income of a ZIP code seems to be associated with a `r get_regression_table(income_model, conf.level = 0.95, digits=2)[[41,3]]` standard-deviation change in the price that consumers in that ZIP code expect to pay for the same product.

# Problem 3
Statement A is **true.** 

Statement B is **undecidable.**

Statement C is **true.**

Statement D is **false.**

Statement E is **true.**